---
title: "Pleural cfDNA Screening Array"
subtitle: "DNAm Data Release Summary"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = TRUE, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("dnam-release-summary.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Overview

This document summarises the versioned DNAm data release for the pleural cfDNA screening array study (project M2644). It links array locations to the released samplesheet, constructs a combined phenotype table annotating each sample with sex, malignancy status, plate, and QC-pass/fail status, and produces a stratified descriptive table to characterise the study population.

```{r globals, results = 'hide', echo = F}
```

```{r files, echo = F}
```

```{r get.samplesheets, echo = F}
```

```{r get.pheno, echo = F}
```

# Sample Characteristics of QC Pass/Fail Samples

Of `r nrow(pheno)` samples, `r sum(pheno$qc.pass == 0)` (`r round(mean(pheno$qc.pass == 0) * 100, 1)`%) failed QC and were excluded from the clean release. The table below compares the distribution of malignancy status, sex, plate, and slide between samples that passed and failed QC, giving an overview of the study population and highlighting any systematic differences between included and excluded samples.

```{r tab.pheno, echo = F}
```
