---
title: "Pleural cfDNA Screening Array"
subtitle: "EWAS summary"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = FALSE, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("ewas-summary.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

```{r globals, results = 'hide', echo = F}
```

```{r load.data, echo = F}
```

```{r source-models, results = 'hide', echo = F}
```

# Population description

## Sample characteristics table

```{r tableone}
```

### Variable defintions
- **meso_malig:** meso v. any other malignancy
- **meso_bape:** meso v. bape
- **epithelioid_other_meso:** epithelioid v. any other type of meso
- **lung_other_malig:** lung vs. any other malignancy
- **malig_non_infection:** malignant vs. non-malignant (excluding infection)
- **infection:** infection vs. non-infection
- **infection_non_malig:** infection vs. non-infection (excluding malignant)

## Relationship between observed and DNAm estimated variables

### Correlation between age and DNAm-age

```{r age.cor}
```

# EWAS 

## models run

```{r models}
```

## Count of significant hits

Number of significant hits identified per model. Statistical significance was determined using a Bonferroni multiple testing threshold correcting for `r nrow(ret$malignant$ret$table)` tests. That is $$0.05/`r nrow(ret$malignant$ret$table)` = `r sprintf("%.2e",0.05/nrow(ret$malignant$ret$table))`$$

Please refer to specific EWAS reports for tables summarising top results. 

```{r hits}
```

## QQ plots

```{r qq}
```

## Manhattan plots

```{r manhattan}
```

## Volcano plots

```{r volcano, fig.width = 14, fig.height = 20}
```

## OSCA plots

Percent of phenotype variance explained by 'omic' measure

```{r osca, fig.width = 15}
```